Financial Contributions to Presidential Campaigns by CA in 2016
========================================================

> **Tip**: One of the requirements of this project is that your code follows
good formatting techniques, including limiting your lines to 80 characters or
less. If you're using RStudio, go into Preferences \> Code \> Display to set up
a margin line to help you keep track of this guideline!

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(lubridate)
library(dplyr)
library(zipcode)
library(ggmap)
getwd()
setwd("D:/Udacity-DAND/project 4")
```

```{r echo=FALSE, Load_the_Data}
original_financial <- read.csv('./P00000001-CA.csv')
```

The dataset was a collection of financial contributions to presidential campaigns by CA in 2016. It contained 19 variables and 130,4346 observations in this dataset. And most variables are factor variables. 

# Univariate Plots Section

## receipt amount distribution
```{r echo=FALSE, Univariate_Plots}
# For CONTRIBUTION RECEIPT AMOUNT, there are many negative values, weird
# Also, I don't know the unit, I think at least it's thousands..
negative_contb_subset <- subset(original_financial, contb_receipt_amt <= 0)
length(negative_contb_subset) #16366 observations with contribution <= 0??

financial <- subset(original_financial, contb_receipt_amt > 0)

ggplot(aes(x = contb_receipt_amt), data = financial) + 
  geom_histogram(bins = 50) + scale_x_log10() +
  labs(x = "contribution receipt amount",
       y = "frequency", title = "Financial Contribution Amount Distribution")
  
# kind like a normal distribution in log scale,tens or hundreds
```

## committee & candidate
```{r echo=FALSE}
length(unique(financial$cmte_id)) # 25 unique committees
length(unique(financial$cand_nm)) # 25 unique candidates
# how to check committee id and candidate id are related one by one?
financial$cmte_cand <- paste(financial$cmte_id, financial$cand_nm, sep = '_')
length(unique(financial$cmte_cand)) #25
# So each candidate has a committee, thus I would focus on the candidate

ggplot(aes(x = cand_nm), data = financial) + geom_bar() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("candidate name") + ylab("contribution times for a candidate") +
  labs(title = "Financial support times for each candidate in CA, 2016")


# This bar chart shows that there were orders of magnitude difference of the financial support times. Two candidates(Clinton, Hillary Rodham and Sanders, Bernard) received financial support for most times. And Hillary Clinton was the top 1 candidate in receiving financial support. Since they are both Democratic, indicating that Democratic is more popular within CA.

# How many candidates are contributed for the general election?
general_election_subset <- subset(financial, election_tp == "G2016")
length(unique(general_election_subset$cand_nm))
# 16 unique candidates were included in the general_election_subset dataset, but also a few candidates didn't receive financial support at all in this period. Only 7 candidates received financial support in general election in CA.

```

## contributing cities
```{r echo=FALSE}
length(unique(financial$contbr_city)) # 2517 unique contributor cities
# head(sort(table(financial$contbr_city), decreasing = TRUE), 10)
top_10_contributing_cities_by_times <- financial %>%
  group_by(contbr_city) %>%
  summarise(n = n()) %>%
  filter(rank(desc(n)) <= 10)

# Maybe I should get information about the city's financial level and their rank
# I want to find the 10 cities with the contribution for 
# quite many times(here not for money)
ggplot(aes(x = contbr_city, y = n), 
       data = top_10_contributing_cities_by_times) +
  geom_histogram(stat = "identity") +
  xlab("city name") + ylab("contribution times") +
  labs(title = "Contribution Times for Top 10 Cities ") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
# We can easily figure out that Los Angeles, San Francisco and San Diego are top 3 contributing cities ranked by contribution times.

```
## occupation
```{r echo=FALSE}
# table(financial$contbr_occupation)
top_10_occupation_by_times <- financial %>%
  group_by(contbr_occupation) %>%
  summarise(n = n()) %>%
  filter(rank(desc(n)) <= 10)

# str(top_10_occupation_by_times)

ggplot(aes(x = contbr_occupation, y = n), 
       data = top_10_occupation_by_times) +
  geom_histogram(stat = "identity") +
  xlab("occupation names") + ylab("contribution times") +
  labs(title = "Contribution Times for Top 10 Occupations ") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
#Actually, I'm quite curious why the retired people contributed most.

```



## contribution time distribution
```{r echo=FALSE}
# wonder if there is a time contributed much larger than any other time,
# and it's related to the actual election event

# Here dt is considered as a factor variable, I need to transform it to datatime first
financial$contb_receipt_dt <- dmy(financial$contb_receipt_dt)
# head(financial$contb_receipt_dt)
# Second challenge: transform the time variable
# Why as.Date and strptime don't work as I expected, they only generated NA
# So I used a package called lubridate, very easy to use
# Now I can group into month
# length(unique(year(financial$contb_receipt_dt)))
# table(year(financial$contb_receipt_dt))
# Wow, the contribution even start on 2013, 
# maybe there will be multiple receipts from same person
financial$contb_receipt_dt_yr_mo <- strftime(financial$contb_receipt_dt, "%Y%m")
# head(financial$contb_receipt_dt_yr_mo)
ggplot(aes(x = contb_receipt_dt_yr_mo), data = financial) + 
  geom_bar() + labs(x = "time", y = "frequency",
                    title = "Financial Contribution Throuth Time") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Actually, we can see most contributions were made in 2016. And there are two peaks , one in 201603(during primary election), and the other in 201610 (one month before the general election)
```

## form type
```{r echo=FALSE}
ggplot(aes(x = form_tp), data = financial) + geom_bar() +
  labs(x = "from type", y = "frequency", title = "Three different form types")
# SA17A£º individual contributions
# SA18: transfers from authorized committees, does it mean negative numbers?
# SA28A: refunds to individuals'
# Still confused on what's going on here
`
Show in New WindowClear OutputExpand/Collapse Output

Show in New WindowClear OutputExpand/Collapse Output
Modify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output
Modify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current ChunkModify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output
Modify Chunk OptionsRun All Chunks AboveRun Current Chunk
Show in New WindowClear OutputExpand/Collapse Output

Show in New WindowClear OutputExpand/Collapse Output

Show in New WindowClear OutputExpand/Collapse Output

Show in New WindowClear OutputExpand/Collapse Output

Show in New WindowClear OutputExpand/Collapse Output
``

## election type
```{r echo=FALSE}
ggplot(aes(x = election_tp), data = financial) + geom_bar() +
  labs(x = "election type", y = "frequency", title = "Financial Contributions for Different Election Types")

# Actullay, most contributions were made for primary election and general election.
```

# Univariate Analysis

### Your idea?

(1) I'm interested to find how was each candidate being supported financially in presidential campign in CA and want to figure out who was the most popular candidate in CA, 2016.
(2) I'm also interested in using financial support data to predict the proportion of votes in general election in CA£¬ 2016. 

### What is the structure of your dataset?

And most of the variables are factor variables, like contributing city, contributing name and etc. It's not a cross section dataset(snapshot taken at a given time). It recorded every financial contribution during the election process.

### What is/are the main feature(s) of interest in your dataset?
- The financial contribution amount distribution is kind of a normal distribution in log scale. And most contributions' amounts are between tens to hundreds.
- There were 25 candidateds being financially supported during the presidential campaign in 2016. The Democratic was well supported compared to the other parties. And Clinton, Hillary Rodham and Sanders, Bernard received top 2 financial contributions ranked by frequency.
- There were more than 2500 cities contributed in this presidential campaign in CA in 2016, of which  Los Angeles, San Francisco and San Diego were top 3 contributing cities ranked by contribution times.
- It was really surprised to find that retired men and not employed people contributed for most times among all occupations.
- When looking at the total contribution amount for each month, we could figure out that most contributions were made in 2016. And there were two contribution peaks, one in 201603(during primary election), and the other in 201610(one month before the general election).
- Actullay, most contributions were made for primary election and general election.
- Most contributions came from indivials, and a few were transferring from authorized committees.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

??can't understand

### Did you create any new variables from existing variables in the dataset?
There are a few changes I made to the original dataset.
(1) I changed the format of the contribution date and also creates another column(called `contb_receipt_dt_yr_mo`) representing the year and month.
(2) I combined the committee id and candidate name to from a new column to check whether committees and candidates were correlated one by one.


### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

(1) There were some records with negative contribution receipt amount in the dataset(original_financial). So I excluded all negative values in mny dataset and I used modified dataset(financial) for my exploration.
(2) And retired man contributed most times of all occupations, which really surprised me.



# Bivariate Plots Section

### candidate & receipt amount
When doing the univariate visualizations and analysis, the y axis for bar charts or histograms is the frequency, or contribution times. And since each contribution receipt amount is different, so the contribution amount distributions differ.

From the first plot, we could see that there was orders of magnitude difference for each candidate. So I changed to log10 scale for better understanding. Of all candidates, Clinton, Hillary Rodham received the most total contribution amount.

```{r echo=FALSE, Bivariate_Plots}
financial.cand_id_amt <- financial %>%
  group_by(cand_nm) %>%
  summarise(total_amt = sum(contb_receipt_amt)) %>%
  arrange(cand_nm)
#head(financial.cand_id_amt)
#str(financial.cand_id_amt)
candidate_p1 <- ggplot(aes(x = cand_nm, y = total_amt), data = financial.cand_id_amt) +
  geom_bar(stat = "identity") +
  labs(x = "candidate name", y = "total amount received",
      title = "Total Financial Support Amount for each candidate") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
candidate_p1
candidate_p2 <- candidate_p1 + scale_y_log10()
candidate_p2

# should I use log10 scale here??
# Who received the most amount?
# financial.cand_id_amt[which.max(financial.cand_id_amt$total_amt),]
# max sum is 93681171 and Clinton, Hillary Rodham receives it

```

### city & receipt amount
Since there were more than 2500 cities in this dataset, so I would still explore top 10 cities, but this time with total contribution amount. While considering the total financial support amount, rather than contribution times, San Francisco and Los Angeles were still top 2 cities, while San Diego was close to other cities.

```{r echo=FALSE}
financial.city_amt <- financial %>%
  group_by(contbr_city) %>%
  summarise(total_amt = sum(contb_receipt_amt)) %>%
  arrange(contbr_city)

# Why there are so many invalid city names???
financial.city_amt[which.max(financial.city_amt$total_amt),]
# Can I get top 10 contribution cities in CA?
o <- order(financial.city_amt$total_amt, decreasing=TRUE)[1:10]
top_10_contributing_cities <- financial.city_amt[o,] 

ggplot(aes(x = contbr_city, y = total_amt ), 
       data = top_10_contributing_cities) + 
  geom_bar(stat = 'identity') +
  labs(x = "city name", y = "total amount received",
      title = "Total Financial Support Amount from Top 10 Cities") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```
### top 10 contribution city & candidate
Here I only cared about three candidates, Clinton, Hillary Rodham, Sanders, Bernard and Trump, Donald J.
```{r}
# Now this is still a bivariate analysis.
# names(financial)
# summary(financial$cand_nm)
top_3_candidates <- c("Clinton, Hillary Rodham", "Sanders, Bernard", "Trump, Donald J.")

# names(top_10_contributing_cities_by_times)
ggplot(aes(x = contbr_city), data = subset(financial, 
      contbr_city %in% top_10_contributing_cities_by_times$contbr_city & cand_nm %in% top_3_candidates )) + geom_bar() +
  facet_wrap(~cand_nm, scales = "free_y") + 
  labs(x = "city name", y = "frequency",
      title = "Total Financial Contribution Times from Top 10 Cities") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
# While the big cities are still the big cities
```

### contribution time & receipt amount

Obviously, 201607-201610 are the most intense periods. While the trend is different from frequency.
```{r echo=FALSE}

financial.dt_amt <- financial %>%
  group_by(contb_receipt_dt_yr_mo) %>%
  summarise(total_amt = sum(contb_receipt_amt)) %>%
  arrange(contb_receipt_dt_yr_mo)

# head(financial.dt_amt)

financial.dt_amt[which.max(financial.dt_amt$total_amt),]
# Which month received most contributions? Is there a cyclic routine?
# like a candidate goes here for a speech and then contribution goes up
ggplot(aes(x = contb_receipt_dt_yr_mo, y = total_amt ), 
       data = financial.dt_amt) + 
  geom_bar(stat = 'identity') +
  labs(x = "time", y = "total amount received",
      title = "Total Financial Support Amount Through Time") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### Financial Contribution Times for Each Candidate During Different Election Period
```{r echo=FALSE}
ggplot(aes(x = cand_nm), data = 
      subset(financial, election_tp == 'G2016' | election_tp == 'P2016')) + 
  geom_bar() +
  facet_wrap(~election_tp, ncol = 2) + 
  labs(x = "candidate name", y = "frequency",
      title = "Financial Contribution Times for Each Candidate\n
      During Different Election Period") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
# We can see that during primary election, more than two candidates were financially supported, which was consistent with the election rule.

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

When doing univariate analysis, I explored the frequency(contribution times) of some variables, like contribution city, candidate name. And when doing bivariate analysis, I primarily changed the frequency(contribution times) to contribution receipt amount. And the most conclusions in univariate analysis didn't vary much, but the distribution changed a little. Like

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?

Since there were only one numerical variable(i.e., contribution receipt amount), so scatterplot was not suitable here. Thus no linear or non-linear relationship were found in this dataset. But We could still see that the contribution receipt amount varied as the candidate or contributing city changed.


# Multivariate Plots Section

### top 10 contribution city & candidate & contribution amount
Here I only cared about three candidates, Clinton, Hillary Rodham, Sanders, Bernard and Trump, Donald J. It's very similar to the frequency one.
```{r}
# Now this is still a bivariate analysis.
# names(financial)
# summary(financial$cand_nm)
top_3_candidates <- c("Clinton, Hillary Rodham", "Sanders, Bernard", "Trump, Donald J.")

# names(top_10_contributing_cities_by_times)
# Should I group_by two variables?
financial.city_cand_amt <- financial %>%
  group_by(cand_nm, contbr_city) %>%
  summarise(total_amt = sum(contb_receipt_amt))
# Do I need to ungroup data??

ggplot(aes(x = contbr_city, y = total_amt), 
       data = subset(financial.city_cand_amt, 
      contbr_city %in% top_10_contributing_cities_by_times$contbr_city & cand_nm %in% top_3_candidates)) + 
  geom_histogram(stat = "identity") +
  facet_wrap(~cand_nm, scales = "free_y") + 
  labs(x = "city name", y = "financial contribution amount",
      title = "Total Financial Support Amount from Top 10 Cities") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
# While the big cities are still the big cities
# I should choose free scale
```

### eleltion type & total contribution amount & candidate
```{r echo=FALSE}
financial.election_tp_cand_amt <- financial %>%
  group_by(cand_nm, election_tp) %>%
  summarise(total_amt = sum(contb_receipt_amt))

ggplot(aes(x = cand_nm, y = total_amt), 
       data = subset(financial.election_tp_cand_amt, 
      election_tp == 'G2016' | election_tp == 'P2016')) + 
  geom_histogram(stat = "identity") +
  facet_wrap(~election_tp, scales = "free_y") + 
  labs(x = "candidate name", y = "financial contribution amount",
      title = "Financial Contribution Times for Each Candidate \n During Different Election Period") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



```

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

### contribution amount & candidate & time
```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x = contb_receipt_dt_yr_mo, y = contb_receipt_amt),
       data = financial) + geom_line() +
  facet_wrap(~cand_nm) + 
  labs(x = "time", y = "contribution receipt amount",
      title = "Financial Support Amount Through Time for Each Candidate") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
### contribution amount & city & time
```{r echo=FALSE}

ggplot(aes(x = contb_receipt_dt_yr_mo, y = contb_receipt_amt),
       data = subset(financial, 
          contbr_city %in% top_10_contributing_cities$contbr_city)) + 
  geom_line() +
  facet_wrap(~contbr_city) + 
  labs(x = "time", y = "contribution receipt amount",
      title = "Financial Support Amount Through Time from Each City") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
### Map distribution
Maybe I can group it by candidate, so
```{r echo=FALSE}
data(zipcode)

financial$modified_contbr_zip <- clean.zipcodes(financial$contbr_zip)
# head(financial$modified_contbr_zip)
# I need to aggregate the same zipcode together, 
# I can first aggregate, then merge it with zipcode
financial.zip_amt <- financial %>%
  group_by(modified_contbr_zip) %>%
  summarise(total_amt = sum(contb_receipt_amt)) %>%
  arrange(modified_contbr_zip)
# str(financial.zip_amt)

new_financial <- merge(financial.zip_amt, zipcode, 
                       by.x='modified_contbr_zip',
                       by.y='zip')
# head(new_financial)
# length(unique(new_financial$modified_contbr_zip)) #2128
# table(new_financial$state) 
# Why there are a few contributions mapped to other state?
# Because they are in different state when they contributed??
# without map version
ggplot(data = subset(new_financial, state == 'CA')) + 
  geom_point(aes(x=longitude, y=latitude, size = total_amt),
             alpha = 1 / 4, color = 'orange')

# with map version
qmplot(longitude, latitude, size = total_amt,
       data = subset(new_financial, state == 'CA'), 
      color = I("red"), alpha = 1 / 20)
# still needs some adjustment
# maybe it's better to use color rather than size to 
# represent total_amt difference
```
### general election prediction
```{r}
#names(financial)
financial_for_general_election <- subset(financial,
          election_tp == 'G2016' & contb_receipt_amt > 0)
financial_for_general_election_sum <- financial_for_general_election %>%
  group_by(cand_nm) %>%
  summarise(total_amt = sum(contb_receipt_amt),
            n = n())
#head(financial_for_general_election_sum)
#financial_for_general_election_sum

financial_for_general_election_sum$amt_ratio <- financial_for_general_election_sum$total_amt / sum(financial_for_general_election_sum$total_amt)

financial_for_general_election_sum$times_ratio <- financial_for_general_election_sum$n / sum(financial_for_general_election_sum$n)
# why above didn't caculate right??
428362 / sum(financial_for_general_election_sum$n)
# financial_for_general_election_sum
# If we use financial support data for simple predicting, we may get the result that Hillary Clinton would win CA, but the actual ratio(61.73%) was much lower than predicted ones. It may be due to bias(because only rich people would contribute to the candidates and normal people were not considred in this prediction, non-respondents bias)
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!